---
title: "Data_Preparation"
author: "Laura Juvany"
date: '2022-11-18'
output: html_document
---

# Data preparation

Upload data of total wet weight per subplot, per site and merge in one data frame

```{r}
#SB with data on cover, height and wet weight
WP1_SB_cover_bm_data <- read.csv("...", sep=";", dec = ",",  header = TRUE)

#HB cover height and wet weight
WP1_HB_cover_bm_data <- read.csv("...", sep=";", dec = ",", header = TRUE)

#NB cover height and wet weight
WP1_NB_cover_bm_data <- read.csv("...", sep=";", dec = ",", header = TRUE)

#join all 3 sites in one data set
WP1_AllSites_cover_bm_data <- rbind(WP1_SB_cover_bm_data, WP1_HB_cover_bm_data, WP1_NB_cover_bm_data)

```

Turn variables from the data sets into numeric

```{r}
WP1_AllSites_cover_bm_data$Wet.weight <- as.numeric(WP1_AllSites_cover_bm_data$Wet.weight)
WP1_AllSites_cover_bm_data$Cover <- as.numeric(WP1_AllSites_cover_bm_data$Cover)
WP1_AllSites_cover_bm_data$Height <- as.numeric(WP1_AllSites_cover_bm_data$Height)
WP1_AllSites_cover_bm_data$Mean.age <- as.numeric(WP1_AllSites_cover_bm_data$Mean.age)

```

Upload data from subsamples used to determine wet to dry biomass

```{r}

#SUBSAMPLE A from All sites data, wet weight and dry weight per plot
WP1_SubA_WetDryWeight <- read.csv("C://Users/lajs0002/Desktop/Laura/WP1/Data_analysis_WP1/WP1_SubA_weights.csv", sep=";", dec = ",", header = TRUE)

```

Turn variables into numeric

```{r}
WP1_SubA_WetDryWeight$Wet.weight.sub.A <- as.numeric(WP1_SubA_WetDryWeight$Wet.weight.sub.A)
WP1_SubA_WetDryWeight$Dry.weight.Sub.A <- as.numeric(WP1_SubA_WetDryWeight$Dry.weight.Sub.A)
```

### Data preparation to obtain the dry biomass for all subplots

Create the variable dry content ratio by dividing the dry weight by the wet weight of each sample

```{r}
WP1_SubA_WetDryWeight$DryBmRatio <- (WP1_SubA_WetDryWeight$Dry.weight.Sub.A/WP1_SubA_WetDryWeight$Wet.weight.sub.A)
```

Handle some problems in the data

```{r}
#Two values that had a typo in the dataset when collecting the data
WP1_SubA_WetDryWeight$DryBmRatio[42] <- 0.4381577
WP1_SubA_WetDryWeight$DryBmRatio[27] <- 0.4362405

#Put values of the ratio of water content to those subsamples with missing values of wet weight or dry weight
WP1_SubA_WetDryWeight$DryBmRatio[14] <- 0.4477928
WP1_SubA_WetDryWeight$DryBmRatio[15] <- 0.36985745
WP1_SubA_WetDryWeight$DryBmRatio[16] <- 0.4104677
WP1_SubA_WetDryWeight$DryBmRatio[55] <- 0.4365602
```

Convert wet weights from all subplots to dry weights with the dry weight ratio from subsamples A

Create a new column in the df where the dry weights of the subsamples will go, and fill in with NA's

```{r}
WP1_AllSites_cover_bm_data$drywt<-NA
```

Create a loop with same length as the created variable, and multiply the wet weight to the Dry biomass ratio with the coinciding Site, Stand, Plot and Species. Use function "try" to make the loop work even if there are NA in the data set (will give errors but the loop works)

```{r}
for(i in 1:length(WP1_AllSites_cover_bm_data$drywt)){
  try(WP1_AllSites_cover_bm_data$drywt[i]<-WP1_AllSites_cover_bm_data$Wet.weight[i]*
        WP1_SubA_WetDryWeight$DryBmRatio[
          WP1_SubA_WetDryWeight$Site==WP1_AllSites_cover_bm_data$Site[i] &
            WP1_SubA_WetDryWeight$Stand==WP1_AllSites_cover_bm_data$Stand[i] &
            WP1_SubA_WetDryWeight$Plot==WP1_AllSites_cover_bm_data$Plot[i] &
            WP1_SubA_WetDryWeight$Species==WP1_AllSites_cover_bm_data$Species[i]])
}

#There were some NA values as there was no biomass Ratio to apply, here I repeat the loop for those NA values and take the ratio from the samples with same species from same stand and site
for(i in which(is.na(WP1_AllSites_cover_bm_data$drywt))){
  WP1_AllSites_cover_bm_data$drywt[i]<-WP1_AllSites_cover_bm_data$Wet.weight[i]*
    mean(WP1_SubA_WetDryWeight$DryBmRatio[
      WP1_SubA_WetDryWeight$Site==WP1_AllSites_cover_bm_data$Site[i] &
        WP1_SubA_WetDryWeight$Stand==WP1_AllSites_cover_bm_data$Stand[i] &
        WP1_SubA_WetDryWeight$Species==WP1_AllSites_cover_bm_data$Species[i]])
}

#There were some NA values as there was no Bm Ratio to apply, here I repeat the loop for those NA values and take the ratio from the samples with same species from same site
for(i in which(is.na(WP1_AllSites_cover_bm_data$drywt))){
  WP1_AllSites_cover_bm_data$drywt[i]<-WP1_AllSites_cover_bm_data$Wet.weight[i]*
    mean(WP1_SubA_WetDryWeight$DryBmRatio[
      WP1_SubA_WetDryWeight$Site==WP1_AllSites_cover_bm_data$Site[i] &
        WP1_SubA_WetDryWeight$Species==WP1_AllSites_cover_bm_data$Species[i]])
}
```
Save the data in a CSV
```{r}
write.csv(WP1_AllSites_cover_bm_data,"...")
```
Add other measured factors in the field in the excel file and upload the data
```{r}
WP1_AllSitesData <- read.csv("...", sep=",", dec = ".", header = TRUE)

```
























