---
title: "DataPreparation_Fractions"
author: "Laura Juvany"
date: '2022-11-18'
output: html_document
---
# Data preparation for the forage biomass and fractions models
Install and call needed packages
```{r}
install.packages("dplyr")
install.packages("tibble")
install.packages("readr")

library(dplyr)
library(tibble)
library(readr)
```
Upload the data from subsample B collected in 2019. These subsamples show the biomass proportion along 7 stem categories, and a sub sample of these was analysed in order to get the proportion of leaves in these 7 stem categories
```{r}
SingPlant_Fractions <- read.csv(".../WeightsSinglePlantSubB_All.csv", sep=",", dec = ".", header = TRUE)
```
Calculate first the proportion of leaves across the samples within the same stand, and later these with the total biomass collected in each plot
Create proportions per each fraction
```{r}
SingPlant_Fractions$Prop_FR1Leaves <- ((((SingPlant_Fractions$FR1_LeavesWeight*100)/SingPlant_Fractions$Leaves_TotalWeight))*0.01)
SingPlant_Fractions$Prop_FR2Leaves <- ((((SingPlant_Fractions$FR2_LeavesWeight*100)/SingPlant_Fractions$Leaves_TotalWeight))*0.01)
SingPlant_Fractions$Prop_FR3Leaves <- ((((SingPlant_Fractions$FR3_LeavesWeight*100)/SingPlant_Fractions$Leaves_TotalWeight))*0.01)
SingPlant_Fractions$Prop_FR4Leaves <- ((((SingPlant_Fractions$FR4_LeavesWeight*100)/SingPlant_Fractions$Leaves_TotalWeight))*0.01)
SingPlant_Fractions$Prop_FR5Leaves <- ((((SingPlant_Fractions$FR5_LeavesWeight*100)/SingPlant_Fractions$Leaves_TotalWeight))*0.01)
SingPlant_Fractions$Prop_FR6Leaves <- ((((SingPlant_Fractions$FR6_LeavesWeight*100)/SingPlant_Fractions$Leaves_TotalWeight))*0.01)
SingPlant_Fractions$Prop_FR7Leaves <- ((((SingPlant_Fractions$FR7_LeavesWeight*100)/SingPlant_Fractions$Leaves_TotalWeight))*0.01)
```
Now, we need to apply the calculated biomass leaf fractions to the total subsample B biomass leaves
First, upload all the data for subsample B, with the weights per stem fraction i the total leaf biomass
```{r}
WP1_SubB <- read.csv("C://Users/lajs0002/Desktop/Laura/WP1/Data_analysis_WP1/BiomassFractions/WeightsSubB_All.csv", sep=",", dec = ".", header = TRUE)
```
Now, apply the leaf ratios across fractions to the fraction data
Create a new column where the dry weights of the leaves will go, and fill it with NA's
```{r}
WP1_SubB$FR1_Leaves<-NA
```
Create a loop with same length as the created variable, and multiply the leaves weight to the leaves biomass ratio per fraction with the coinciding Site, Stand, (Plot) and Species
Use the function try to make the loop work even if there are NA in the data set (will give errors but the loop works)
```{r}
#For fraction 1
for(i in 1:length(WP1_SubB$FR1_Leaves)){
  try(WP1_SubB$FR1_Leaves[i]<-WP1_SubB$FRLeaves_weight[i]*
        SingPlant_Fractions$Prop_FR1Leaves[
          SingPlant_Fractions$Site==WP1_SubB$Site[i] &
            SingPlant_Fractions$Stand==WP1_SubB$Stand[i] &
            SingPlant_Fractions$Species==WP1_SubB$Species[i]])
}

#For fraction 2
WP1_SubB$FR2_Leaves<-NA

for(i in 1:length(WP1_SubB$FR2_Leaves)){
  try(WP1_SubB$FR2_Leaves[i]<-WP1_SubB$FRLeaves_weight[i]*
        SingPlant_Fractions$Prop_FR2Leaves[
          SingPlant_Fractions$Site==WP1_SubB$Site[i] &
            SingPlant_Fractions$Stand==WP1_SubB$Stand[i] &
            SingPlant_Fractions$Species==WP1_SubB$Species[i]])
}

#For fraction 3
WP1_SubB$FR3_Leaves<-NA

for(i in 1:length(WP1_SubB$FR3_Leaves)){
  try(WP1_SubB$FR3_Leaves[i]<-WP1_SubB$FRLeaves_weight[i]*
        SingPlant_Fractions$Prop_FR3Leaves[
          SingPlant_Fractions$Site==WP1_SubB$Site[i] &
            SingPlant_Fractions$Stand==WP1_SubB$Stand[i] &
            SingPlant_Fractions$Species==WP1_SubB$Species[i]])
}

#For fraction 4
WP1_SubB$FR4_Leaves<-NA

for(i in 1:length(WP1_SubB$FR4_Leaves)){
  try(WP1_SubB$FR4_Leaves[i]<-WP1_SubB$FRLeaves_weight[i]*
        SingPlant_Fractions$Prop_FR4Leaves[
          SingPlant_Fractions$Site==WP1_SubB$Site[i] &
            SingPlant_Fractions$Stand==WP1_SubB$Stand[i] &
            SingPlant_Fractions$Species==WP1_SubB$Species[i]])
}

#For fraction 5
WP1_SubB$FR5_Leaves<-NA

for(i in 1:length(WP1_SubB$FR5_Leaves)){
  try(WP1_SubB$FR5_Leaves[i]<-WP1_SubB$FRLeaves_weight[i]*
        SingPlant_Fractions$Prop_FR5Leaves[
          SingPlant_Fractions$Site==WP1_SubB$Site[i] &
            SingPlant_Fractions$Stand==WP1_SubB$Stand[i] &
            SingPlant_Fractions$Species==WP1_SubB$Species[i]])
}

#For fraction 6
WP1_SubB$FR6_Leaves<-NA

for(i in 1:length(WP1_SubB$FR6_Leaves)){
  try(WP1_SubB$FR6_Leaves[i]<-WP1_SubB$FRLeaves_weight[i]*
        SingPlant_Fractions$Prop_FR6Leaves[
          SingPlant_Fractions$Site==WP1_SubB$Site[i] &
            SingPlant_Fractions$Stand==WP1_SubB$Stand[i] &
            SingPlant_Fractions$Species==WP1_SubB$Species[i]])
}

#For fraction 7
WP1_SubB$FR7_Leaves<-NA

for(i in 1:length(WP1_SubB$FR7_Leaves)){
  try(WP1_SubB$FR7_Leaves[i]<-WP1_SubB$FRLeaves_weight[i]*
        SingPlant_Fractions$Prop_FR7Leaves[
          SingPlant_Fractions$Site==WP1_SubB$Site[i] &
            SingPlant_Fractions$Stand==WP1_SubB$Stand[i] &
            SingPlant_Fractions$Species==WP1_SubB$Species[i]])
}
```
Check if it worked by adding up all the leaves biomass fractions from the Subsamples B
```{r}
WP1_SubB$Check_SumLeaves <- (WP1_SubB$FR1_Leaves + WP1_SubB$FR2_Leaves + WP1_SubB$FR3_Leaves + WP1_SubB$FR4_Leaves + WP1_SubB$FR5_Leaves + WP1_SubB$FR6_Leaves + WP1_SubB$FR7_Leaves)
```
Download the data into a CSV
```{r}
write.csv(WP1_SubB,"C:/Users/lajs0002/Desktop/Laura/WP1/Data_analysis_WP1/BiomassFractions/WP1_SubB_StemLeaves.csv")
```
Now, we need to make some small changes to the data
1. Download the data and add in the excel the data from those samples that were so small, that the whole sample was processed as single plant analysis, so the data is also included in this data set 
These samples are:
SB: sample 6, 11, 17, 29, 71
HB: sample 4, 67, 78
NB: sample 33
2. Also, we decided that for those samples where the single analysed model plant had a larger number of stem fractions than the whole sample from a plot, and these had leaf biomass for those fractions and the larger plot sample did not, we decided to add that weight to the next smaller leaf fraction for that sample.
EXAMPLE: Stand 3 model plant for bilberry has stem fractions until 7 and leaves until FR6, but the samples from Stand 3 Plot 2 have only stem fraction until FR5, the loop would assign anyway the proportion of leaves to FR6, but we took this weight and summed it to the leaves fraction in FR5
3. Sample 10 NB leaves fraction was missing on the original data, so I found the sample with all the fractions in the bag and re-weighed them again to get the data.
Now, upload the data with these changes
```{r}
WP1_SubB_WithChanges <- read.csv("C://Users/lajs0002/Desktop/Laura/WP1/Data_analysis_WP1/BiomassFractions/WP1_SubB_StemLeaves_WithChanges_New.csv", sep=",", dec = ".", header = TRUE)
```
Now we need to make the proportions of total fraction weight, stem and leaves separately
Create a new data frame with only the necessary variables, which are: total sample weight, total stem weight and total leaves weight
```{r}
WP1_SubB_CalcProp <- WP1_SubB_WithChanges[, colnames(WP1_SubB_WithChanges)[c(1:13,16:22)]]
```
Calculate total stem weight
```{r}
WP1_SubB_CalcProp <- WP1_SubB_CalcProp %>% rowwise() %>%
  mutate(TotalStemWeight = sum(c_across(FR1_weight:FR7_weight)))
```
Calculate total leaves weight
```{r}
WP1_SubB_CalcProp <- WP1_SubB_CalcProp %>% rowwise() %>%
  mutate(TotalLeavesWeight = sum(c_across(FR1_Leaves:FR7_Leaves)))
```
Calculate total sample weight
```{r}
WP1_SubB_CalcProp$TotalSampleWeight <- (WP1_SubB_CalcProp$TotalStemWeight + WP1_SubB_CalcProp$TotalLeavesWeight)
```
Create proportions
From total weight of the sample, calculate the proportion of each stem fraction
```{r}
WP1_SubB_CalcProp$Prop_FR1Stem <- ((((WP1_SubB_CalcProp$FR1_weight*100)/WP1_SubB_CalcProp$TotalSampleWeight))*0.01)
WP1_SubB_CalcProp$Prop_FR2Stem <- ((((WP1_SubB_CalcProp$FR2_weight*100)/WP1_SubB_CalcProp$TotalSampleWeight))*0.01)
WP1_SubB_CalcProp$Prop_FR3Stem <- ((((WP1_SubB_CalcProp$FR3_weight*100)/WP1_SubB_CalcProp$TotalSampleWeight))*0.01)
WP1_SubB_CalcProp$Prop_FR4Stem <- ((((WP1_SubB_CalcProp$FR4_weight*100)/WP1_SubB_CalcProp$TotalSampleWeight))*0.01)
WP1_SubB_CalcProp$Prop_FR5Stem <- ((((WP1_SubB_CalcProp$FR5_weight*100)/WP1_SubB_CalcProp$TotalSampleWeight))*0.01)
WP1_SubB_CalcProp$Prop_FR6Stem <- ((((WP1_SubB_CalcProp$FR6_weight*100)/WP1_SubB_CalcProp$TotalSampleWeight))*0.01)
WP1_SubB_CalcProp$Prop_FR7Stem <- ((((WP1_SubB_CalcProp$FR7_weight*100)/WP1_SubB_CalcProp$TotalSampleWeight))*0.01)
```
From total weight of the sample, calculate the proportion of leaves per fraction
```{r}
WP1_SubB_CalcProp$Prop_FR1Leaves <- ((((WP1_SubB_CalcProp$FR1_Leaves*100)/WP1_SubB_CalcProp$TotalSampleWeight))*0.01)
WP1_SubB_CalcProp$Prop_FR2Leaves <- ((((WP1_SubB_CalcProp$FR2_Leaves*100)/WP1_SubB_CalcProp$TotalSampleWeight))*0.01)
WP1_SubB_CalcProp$Prop_FR3Leaves <- ((((WP1_SubB_CalcProp$FR3_Leaves*100)/WP1_SubB_CalcProp$TotalSampleWeight))*0.01)
WP1_SubB_CalcProp$Prop_FR4Leaves <- ((((WP1_SubB_CalcProp$FR4_Leaves*100)/WP1_SubB_CalcProp$TotalSampleWeight))*0.01)
WP1_SubB_CalcProp$Prop_FR5Leaves <- ((((WP1_SubB_CalcProp$FR5_Leaves*100)/WP1_SubB_CalcProp$TotalSampleWeight))*0.01)
WP1_SubB_CalcProp$Prop_FR6Leaves <- ((((WP1_SubB_CalcProp$FR6_Leaves*100)/WP1_SubB_CalcProp$TotalSampleWeight))*0.01)
WP1_SubB_CalcProp$Prop_FR7Leaves <- ((((WP1_SubB_CalcProp$FR7_Leaves*100)/WP1_SubB_CalcProp$TotalSampleWeight))*0.01)
```
Check if they add up
```{r}
WP1_SubB_CalcProp <- WP1_SubB_CalcProp %>% rowwise() %>%
  mutate(CheckStemProp = sum(c_across(Prop_FR1Stem:Prop_FR7Stem)))

WP1_SubB_CalcProp <- WP1_SubB_CalcProp %>% rowwise() %>%
  mutate(CheckLeavesProp = sum(c_across(Prop_FR1Leaves:Prop_FR7Leaves)))
```
Add up proportions from leaves and stems from each fraction to analyse later
```{r}
WP1_SubB_CalcProp$Prop_FR1_SandL <- (WP1_SubB_CalcProp$Prop_FR1Stem + WP1_SubB_CalcProp$Prop_FR1Leaves)
WP1_SubB_CalcProp$Prop_FR2_SandL <- (WP1_SubB_CalcProp$Prop_FR2Stem + WP1_SubB_CalcProp$Prop_FR2Leaves)
WP1_SubB_CalcProp$Prop_FR3_SandL <- (WP1_SubB_CalcProp$Prop_FR3Stem + WP1_SubB_CalcProp$Prop_FR3Leaves)
WP1_SubB_CalcProp$Prop_FR4_SandL <- (WP1_SubB_CalcProp$Prop_FR4Stem + WP1_SubB_CalcProp$Prop_FR4Leaves)
WP1_SubB_CalcProp$Prop_FR5_SandL <- (WP1_SubB_CalcProp$Prop_FR5Stem + WP1_SubB_CalcProp$Prop_FR5Leaves)
WP1_SubB_CalcProp$Prop_FR6_SandL <- (WP1_SubB_CalcProp$Prop_FR6Stem + WP1_SubB_CalcProp$Prop_FR6Leaves)
WP1_SubB_CalcProp$Prop_FR7_SandL <- (WP1_SubB_CalcProp$Prop_FR7Stem + WP1_SubB_CalcProp$Prop_FR7Leaves)
```
Here I remove those samples that have NS because we discarded them (& in total) or we couldn't find them (1/2)
```{r}
WP1_SubB_CalcProp <- na.omit(WP1_SubB_CalcProp)
```
Make new data frame with those columns of data I want to analyse
```{r}
WP1_SubB_Prop_SandL <- WP1_SubB_CalcProp[, colnames(WP1_SubB_CalcProp)[c(1:6,38:44)]]
```
Save as CSV
```{r}
write.csv(WP1_SubB_Prop_SandL,".../WP1_SubB_Prop_SandL.csv")
```
Need to add the independent variables.
We will start with the plot level ones, like basal areas, mean age and site index.
These independent variables are are found in the dataframe used to build from cover to biomass models, the file is called "WP1_AllSitesData"
```{r}
WP1_AllSitesData <- read.csv("C://Users/lajs0002/Desktop/Laura/WP1/Data_analysis_WP1/WP1_AllSitesData.csv", sep=",", dec = ".", header = TRUE)

WP1_SubB_Prop_SandL <- read.csv("C:/Users/lajs0002/Desktop/Laura/WP1/Data_analysis_WP1/BiomassFractions/WP1_SubB_Prop_SandL.csv")
```
Will include these with a loop
```{r}
#PlotBA
WP1_SubB_Prop_SandL$PlotBA<-NA
for(i in 1:length(WP1_SubB_Prop_SandL$PlotBA)){
  try(WP1_SubB_Prop_SandL$PlotBA[i]<-WP1_AllSitesData$PlotBA[
            WP1_AllSitesData$Site==WP1_SubB_Prop_SandL$Site[i] &
            WP1_AllSitesData$Stand==WP1_SubB_Prop_SandL$Stand[i] &
            WP1_AllSitesData$Plot==WP1_SubB_Prop_SandL$Plot[i]])
}
#PineBA
WP1_SubB_Prop_SandL$PineBA<-NA
for(i in 1:length(WP1_SubB_Prop_SandL$PineBA)){
  try(WP1_SubB_Prop_SandL$PineBA[i]<-WP1_AllSitesData$PineBA[
    WP1_AllSitesData$Site==WP1_SubB_Prop_SandL$Site[i] &
      WP1_AllSitesData$Stand==WP1_SubB_Prop_SandL$Stand[i] &
      WP1_AllSitesData$Plot==WP1_SubB_Prop_SandL$Plot[i]])
}
#SpruceBA
WP1_SubB_Prop_SandL$SpruceBA<-NA
for(i in 1:length(WP1_SubB_Prop_SandL$SpruceBA)){
  try(WP1_SubB_Prop_SandL$SpruceBA[i]<-WP1_AllSitesData$SpruceBA[
    WP1_AllSitesData$Site==WP1_SubB_Prop_SandL$Site[i] &
      WP1_AllSitesData$Stand==WP1_SubB_Prop_SandL$Stand[i] &
      WP1_AllSitesData$Plot==WP1_SubB_Prop_SandL$Plot[i]])
}
#DeciduousBA
WP1_SubB_Prop_SandL$DeciduousBA<-NA
for(i in 1:length(WP1_SubB_Prop_SandL$DeciduousBA)){
  try(WP1_SubB_Prop_SandL$DeciduousBA[i]<-WP1_AllSitesData$DeciduousBA[
    WP1_AllSitesData$Site==WP1_SubB_Prop_SandL$Site[i] &
      WP1_AllSitesData$Stand==WP1_SubB_Prop_SandL$Stand[i] &
      WP1_AllSitesData$Plot==WP1_SubB_Prop_SandL$Plot[i]])
}
#PlotSI
WP1_SubB_Prop_SandL$StandSI<-NA
for(i in 1:length(WP1_SubB_Prop_SandL$StandSI)){
  try(WP1_SubB_Prop_SandL$StandSI[i]<-WP1_AllSitesData$SI_M1[
    WP1_AllSitesData$Site==WP1_SubB_Prop_SandL$Site[i] &
      WP1_AllSitesData$Stand==WP1_SubB_Prop_SandL$Stand[i] &
      WP1_AllSitesData$Plot==WP1_SubB_Prop_SandL$Plot[i]])
}
#Stand mean age
WP1_SubB_Prop_SandL$MeanAge<-NA
for(i in 1:length(WP1_SubB_Prop_SandL$MeanAge)){
  try(WP1_SubB_Prop_SandL$MeanAge[i]<-WP1_AllSitesData$Mean.age[
    WP1_AllSitesData$Site==WP1_SubB_Prop_SandL$Site[i] &
      WP1_AllSitesData$Stand==WP1_SubB_Prop_SandL$Stand[i] &
      WP1_AllSitesData$Plot==WP1_SubB_Prop_SandL$Plot[i]])
}
```
Now we need to add ramet height and percentage cover per species and make a mean per plot
Create a new df for the mean height and rename its columns
```{r}
WP1_MeanShrubHeight <- (aggregate(WP1_AllSitesData$Height,list(WP1_AllSitesData$Site, WP1_AllSitesData$Stand, WP1_AllSitesData$Plot, WP1_AllSitesData$Species), mean))
nms <- c("Site", "Stand", "Plot", "Species", "MeanHeight")
WP1_MeanShrubHeight <- setNames(WP1_MeanShrubHeight, nms)
```
Now apply these to the large data set
```{r}
WP1_SubB_Prop_SandL$MeanShrubHeight <-NA
for(i in 1:length(WP1_SubB_Prop_SandL$MeanShrubHeight)){
  try(WP1_SubB_Prop_SandL$MeanShrubHeight[i]<-WP1_MeanShrubHeight$MeanHeight[
    WP1_MeanShrubHeight$Site==WP1_SubB_Prop_SandL$Site[i] &
      WP1_MeanShrubHeight$Stand==WP1_SubB_Prop_SandL$Stand[i] &
      WP1_MeanShrubHeight$Plot==WP1_SubB_Prop_SandL$Plot[i] &
      WP1_MeanShrubHeight$Species==WP1_SubB_Prop_SandL$Species[i]])
}
```
o the same for percentage cover
```{r}
WP1_Cover <- (aggregate(WP1_AllSitesData$Cover,list(WP1_AllSitesData$Site, WP1_AllSitesData$Stand, WP1_AllSitesData$Plot, WP1_AllSitesData$Species), mean))

nmr <- c("Site", "Stand", "Plot", "Species", "Cover")
WP1_Cover <- setNames(WP1_Cover, nmr)

WP1_SubB_Prop_SandL$Cover <-NA
for(i in 1:length(WP1_SubB_Prop_SandL$Cover)){
  try(WP1_SubB_Prop_SandL$Cover[i]<-WP1_Cover$Cover[
    WP1_MeanShrubHeight$Site==WP1_SubB_Prop_SandL$Site[i] &
      WP1_MeanShrubHeight$Stand==WP1_SubB_Prop_SandL$Stand[i] &
      WP1_MeanShrubHeight$Plot==WP1_SubB_Prop_SandL$Plot[i] &
      WP1_MeanShrubHeight$Species==WP1_SubB_Prop_SandL$Species[i]])
}
```
Save the data as CSV
```{r}
write.csv(WP1_SubB_Prop_SandL,".../WP1_SubB_Prop_SandL_Competition.csv")
```






















